---
title: "FNDDS to CTO Flow"
author: "Jose Collado"
date: "2026-07-10"
output: html_document
---
### Inital Set up

```{r setup, include=FALSE}
#honestly i forget what this is for but it makes me feel safe when coding
knitr::opts_chunk$set(echo = TRUE)
```

```{r color palettes, echo=FALSE}
#not sure if we'll make any figures but this helps keep color choices accessible
safe_colorblind_palette <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")
```

```{r directory set up, include=FALSE}
# where my files live on my computer. Replace with your directory
setwd("C:/Users/josecoll/Downloads/2026.07.10 FNDDS Files")

# to confirm above change
getwd()
```

## Survey CTO Searchable File

Loading and combining the files from the FNDDS 21-23 documentation page.  
```{r load and clean csv datasets, message=FALSE, warning=FALSE}
library(tidyverse) # general frame work
library(gcplyr) # for read_wides
library(dplyr) # inner_join, piping

# The FNDDS 21-23 file with additional description information
FNDDS_addidesc <- read_wides(files = "FNDDS_addi_desc_raw.csv")

# The FNDDS 21-23 file with base ingredient description
FNDDS_ingre <- read_wides(files = "FNDDS_Ingredients_raw.csv")

# let's merge description files by shared columns. Using full_join retains all rows and inserts NA where data is missing from either file. 'by' let's me specify which columns in the df are the same to lead the join process.
FNDDS_desc_merged <- full_join(FNDDS_addidesc, FNDDS_ingre, by = c("Food_code","Main_food_description", "WWEIA_Category_number","WWEIA_Category_description"))

# Following was to see what the additional row is from. 

#FNDDS_desc_merged %>%
  #group_by(WWEIA_Category_number) %>%
  #count(WWEIA_Category_description)

#FNDDS_ingre %>%
  #group_by(WWEIA_Category_number) %>%
  #count(WWEIA_Category_description)

# After reviewing categories, Human milk was not in the FNDDS ingredients file.

# now remove unnecessary columns (the last three might be important but I would not know atm)
FNDDS_desc_merged$file.x <- NULL
FNDDS_desc_merged$file.y <- NULL
FNDDS_desc_merged$Ingredient_weight_gram <- NULL
FNDDS_desc_merged$Retention_code <- NULL
FNDDS_desc_merged$Moisture_change_percentage <- NULL

# change column types as needed
FNDDS_desc_merged <- FNDDS_desc_merged %>% mutate(across(c(Food_code, WWEIA_Category_number, Seq_num, Ingredient_code), as.numeric))

# check updated column types
str(FNDDS_desc_merged)

# clear unneeded df from global environment
rm(FNDDS_addidesc)
rm(FNDDS_ingre)
```

### Adding WWEIA Categories

Category 1 - Highest level grouping (i.e Milk and Dairy, Protein).
```{r add WWEIA highest level categories}
library(tidyverse) # general frame work
library(gcplyr) # 
library(dplyr) # for coalesce, piping

# load WWEIA category 1 code ranges. Values from FNDDS documentation
WWEIA_category_1_ranges <- read_wides(files = "WWEIA_Category_1_ranges.csv")

# df to hold WWEIA values by row from main df in the loop 
category_holder <- cbind(FNDDS_desc_merged["WWEIA_Category_number"])

# empty column to place values from the for loop
category_holder <- category_holder %>% 
  mutate(WWEIA_Category_1 = NA, .after = 1)

# blank df to fill pulled values in the loop
FNDDS_Merged <- data.frame() 

# loop with subsetting method. (previously did not subset and process took way longer)
for (i in 1:nrow(WWEIA_category_1_ranges)) {
  # load category into variable 
  current_group <- WWEIA_category_1_ranges$Category_1[i]
  
  # second df to hold info. Filters category_holder by WWEIA ranges before adding second column with pasted category name
  FNDDS_holder_2 <- category_holder %>% 
    filter(WWEIA_Category_number >= WWEIA_category_1_ranges$C1_min[i], 
           WWEIA_category_1_ranges$C1_max[i] >= WWEIA_Category_number) %>%
    mutate(WWEIA_Category_1 = paste0(WWEIA_category_1_ranges$Category_1[i]))
  
  # add values from above into this merged file
  FNDDS_Merged <- bind_rows(FNDDS_Merged,FNDDS_holder_2)
}

# add NA column to fill in specific location
FNDDS_desc_merged <- FNDDS_desc_merged %>% 
  mutate(WWEIA_Category_1 = NA, .after = 4)

# replace NA values with newly labeled values
FNDDS_desc_merged$WWEIA_Category_1 <- FNDDS_Merged$WWEIA_Category_1

# clear unneeded df / variables
rm(category_holder)
rm(FNDDS_Merged)
rm(WWEIA_category_1_ranges)
rm(FNDDS_holder_2)
```

Category 2 - Next level grouping (i.e Milk and Dairy > Milk, Flavored Milk, Yogurt).
```{r add WWEIA category 2}
library(tidyverse) # general frame work
library(gcplyr) # 
library(dplyr) # for coalesce

# load WWEIA category 2 code ranges
WWEIA_category_2_ranges <- read_wides(files = "WWEIA_Category_2_ranges.csv")

# holder df to label later
category_holder <- cbind(FNDDS_desc_merged["WWEIA_Category_number"])
category_holder <- category_holder %>% 
  mutate(WWEIA_Category_2 = NA, .after = 1)

#blank df to fill later
FNDDS_Merged <- data.frame() 

for (i in 1:nrow(WWEIA_category_2_ranges)) {
  # load category into variable and print 
  current_group <- WWEIA_category_2_ranges$Category_1[i]
  current_subgroup <- WWEIA_category_2_ranges$Category_2[i]
  
  # second df to hold info. Filters category_holder by WWEIA ranges before adding second column with 
  # pasted category name
  FNDDS_holder_2 <- category_holder %>% filter(WWEIA_Category_number >= WWEIA_category_2_ranges$C2_min[i], WWEIA_category_2_ranges$C2_max[i] >= WWEIA_Category_number) %>%
    mutate(WWEIA_Category_2 = paste0(WWEIA_category_2_ranges$Category_2[i]))
  
  # add values from above into this merged file
  FNDDS_Merged <- bind_rows(FNDDS_Merged,FNDDS_holder_2)
}

# add NA column to fill in specified location
FNDDS_desc_merged <- FNDDS_desc_merged %>% 
  mutate(WWEIA_Category_2 = NA, .after = 5)

# replace NA values with newly labeled values
FNDDS_desc_merged$WWEIA_Category_2 <- FNDDS_Merged$WWEIA_Category_2

# clear unneeded df / variables
rm(category_holder)
rm(FNDDS_Merged)
rm(WWEIA_category_2_ranges)
rm(FNDDS_holder_2)
```


```{r reorganize main dataframe}
library(dplyr) # piping format

# reorganizing the df to make sense to me and hopefully be more readable
FNDDS_desc_merged <- FNDDS_desc_merged %>% relocate(Main_food_description, .after = WWEIA_Category_description) %>%
  relocate(Additional_food_description, .after = Main_food_description) %>%
  relocate(Food_code, .after = Seq_num)

# renaming columns as needed 
FNDDS_desc_merged <- FNDDS_desc_merged %>%
  rename(WWEIA_Category_3 = WWEIA_Category_description)
```

### Adding Survey CTO readable codes

Giving each value a unique numeric value allows Survey CTO to function as intented (I learned this the hard way). After same tinkering, I made it work for each category. At this level, I have also decide to add Condiment and Mixed Dish status markers. Since all of the values are at the category 1 level it made sense for me to do it here.

Category 1 Labeling
```{r Add CTO Category 1 values + Condiment & Mixed Dish}
library(tidyverse) # general
library(dplyr) # piping format

# Table with selection groupings and values in a Survey CTO Readable format
Item_values <- read_wides(files = "SurveyCTO_Item_codes.csv")

# create df with only category 1 values
category_level_values <- Item_values %>%
  filter(Grouping == "Category 1")

# get the number of categories for level 1
row_count <- Item_values %>%
  filter(Grouping == "Category 1") %>%
  nrow()

# holder df to label later with a blank column to add to
Group_value_holder <- cbind(FNDDS_desc_merged["WWEIA_Category_1"])
Group_value_holder <- Group_value_holder %>% 
  mutate(WWEIA_Category_1_code = NA, .after = 1)

# blank data frame to use in the for loop
Item_codes_Merged <- data.frame() 

# loop to go through each level group and add the survey CTO code value
for (i in 1:row_count) {
  # load category into variable
  current_group <- category_level_values$Category[i]
  
  # load category CTO code into variable
  group_code <- category_level_values$Category_key[i]
  
  # load condiment status into variable
  condiment_status <- category_level_values$Condiment_status[i]
  
  # load mix dish status into variable
  mixdish_status <- category_level_values$Mixdish_status[i]
  
  # second df to hold info. Filters Group_value_holder by WWEIA category before adding columns with pasted CTO code value and condiment status 
  Group_value_holder_2 <- Group_value_holder %>% filter(WWEIA_Category_1 == current_group) %>%
    mutate(CTO_code_1 = group_code) %>%
    mutate(Condiment_status = condiment_status) %>%
    mutate(Mixdish_status = mixdish_status) %>%
    select(c(CTO_code_1, Condiment_status, Mixdish_status))
  
  # adds newly saved CTO values to passed df
  Item_codes_Merged <- bind_rows(Item_codes_Merged,Group_value_holder_2)
  
  # prints current group and total row count of passed df
  print(paste0(current_group,":",count(Item_codes_Merged)))
}

# add NA column to fill in specific location
FNDDS_desc_merged <- FNDDS_desc_merged %>% 
  mutate(Category_1_code = NA, .after = 2) %>%
  mutate(Condiment_status = NA, .after = 7) %>%
  mutate(Mixdish_status = NA, .after = 8)

# replace NA values with newly labeled values
FNDDS_desc_merged$Category_1_code <- Item_codes_Merged$CTO_code_1
FNDDS_desc_merged$Condiment_status <- Item_codes_Merged$Condiment_status
FNDDS_desc_merged$Mixdish_status <- Item_codes_Merged$Mixdish_status

# double check values lined up
FNDDS_desc_merged %>%
  group_by(Category_1_code) %>%
  count(WWEIA_Category_1)

# clean up environment
rm(condiment_status)
rm(mixdish_status)
```

Category 2 Labeling
```{r Add CTO Category 2 values}

# create df with only category 2 values
category_level_values <- Item_values %>%
  filter(Grouping == "Category 2")

# get the number of categories for level 2
row_count <- Item_values %>%
  filter(Grouping == "Category 2") %>%
  nrow()

# holder df to label later
Group_value_holder <- cbind(FNDDS_desc_merged["WWEIA_Category_2"])
Group_value_holder <- Group_value_holder %>% 
  mutate(WWEIA_Category_2_code = NA, .after = 1)

# blank data frame to add to
Item_codes_Merged <- data.frame() 

for (i in 1:row_count) {
  # load category into variable
  current_group <- category_level_values$Category[i]
  
  # load category CTO code into variable
  group_code <- category_level_values$Category_key[i]
  
  # second df to hold info. Filters Group_value_holder by WWEIA category before adding second column with pasted CTO code value
  Group_value_holder_2 <- Group_value_holder %>% filter(WWEIA_Category_2 == current_group) %>%
    mutate(CTO_code_2 = group_code) %>%
    select(CTO_code_2)
  
  # adds newly saved CTO values to passed df
  Item_codes_Merged <- bind_rows(Item_codes_Merged,Group_value_holder_2)
  
  # prints current group and total row count of passed df to check
  #print(paste0(current_group,":",count(Item_codes_Merged)))
}

# add NA column to fill in specific location
FNDDS_desc_merged <- FNDDS_desc_merged %>% 
  mutate(Category_2_code = NA, .after = 4)

# replace NA values with newly labeled values
FNDDS_desc_merged$Category_2_code <- Item_codes_Merged$CTO_code_2

# double check values lined up
FNDDS_desc_merged %>%
  group_by(Category_2_code) %>%
  count(WWEIA_Category_2)
```

Category 3 and 4 were already in the FNDDS files as WWEIA_Category_description and Main_food_description. But at these levels, it gets trickier as some groups category 2 and 3 are the same (i.e Cheese). After some tinkering, this was the solution I came up with.
```{r Add CTO Category 3 values}
# Goal: check if the WWEI cat 3 code is the same as the cat 2 code. 

# make a df that has category 2 and 3 values + a blank comparison column
Group_value_holder <- cbind(FNDDS_desc_merged["WWEIA_Category_2"])
Group_value_holder$WWEIA_Category_3 <- FNDDS_desc_merged$WWEIA_Category_3
Group_value_holder$comparison <- NA


# get number of rows in the comparison df
row_count <- Group_value_holder %>%
  nrow()

# compare values of each row. If the same, row = 1, else row = 0
for (i in 1:row_count){
  if (Group_value_holder$WWEIA_Category_2[i] == Group_value_holder$WWEIA_Category_3[i]){
    Group_value_holder$comparison[i] = 0
  } else {
    Group_value_holder$comparison[i] = 1
  }
}

# create df with only category 3 values
category_level_values <- Item_values %>%
  filter(Grouping == "Category 3")

# for each row that is matching (1), look at WWEIA_category_3 and save that as a variable. Use that variable to find the Survey CTO code value and paste the code into the comparison column. For non matching rows (0), move to the next line.

for (i in 1:row_count){
  if (Group_value_holder$comparison[i] == 1){
    current_group <- Group_value_holder$WWEIA_Category_3[i]
    
    Group_value_holder$comparison[i] <- Item_values %>%
      filter(Category == current_group) %>%
      pull(Category_key)
  } else {
    next
  }
}

# add NA column to fill
FNDDS_desc_merged <- FNDDS_desc_merged %>% 
  mutate(Category_3_code = Group_value_holder$comparison, .after = 6)

# replace NA values with newly labeled values
FNDDS_desc_merged$Category_3_code <- Group_value_holder$comparison

# double check values lined up
FNDDS_desc_merged %>%
  group_by(Category_3_code) %>%
  count(WWEIA_Category_3)

# clear the global environment
rm(category_level_values)
rm(Group_value_holder)
rm(Group_value_holder_2)
rm(Item_codes_Merged)
rm(current_group)
rm(current_subgroup)
rm(i)
rm(row_count)
rm(group_code)
```

26.06.16 After months of sidetracks and other tasks that got in the way, I am finally tackling adding the fermentation status variable to add to the master csv file. I will have to reorg this sheet at some point but oh well.

Point of this block is to 1) get high confidence ffbin1 for each unique fndds code and 2) get highest ferm value for each unique fndds code.

```{r Add fermentation status}
library(tidyverse) # general frame work
library(gcplyr) # for read_wides
library(dplyr) # inner_join, piping

# The FNDDS 21-23 file with additional description information
FNDDS_ferm_zpg <- read_wides(files = "fndds23ingfflm_raw.csv")

# check column types
str(FNDDS_ferm_zpg)

# Keep a df with all unique FNDDS codes with FFbin1 (high confidence)
FNDDS_ferm_bin1 <- FNDDS_ferm_zpg %>% 
  group_by(wweaicatnumfood, wweaicatdescfood, parentfooddesc,ffbin1, parentfoodcode) %>%
  distinct(parentfoodcode, ffbin1, .keep_all = TRUE)

# Trying to make a data frame that keeps the highest value of ferming will be more complicated sadly but I will try

# change ferming column to numeric for comparison purposes
FNDDS_ferm_ingcode <- FNDDS_ferm_zpg %>%
  mutate(ingfermscore = as.numeric(ingfermscore)) %>%
  distinct(parentfoodcode, ingfermscore, .keep_all = FALSE)

# count all the unique fndds codes
row_count <- FNDDS_ferm_bin1 %>%
  nrow()

# save a 0 comparison variable for the nested loop
temp <- 0 

#df to append all the values to.
FNDDS_ferm_ingcode_merged <- data.frame(parentfoodcode=character(),
                                        merged_ferming=numeric())

# create a loop to 1) subset df by fndds code, 2) compare ferming code and keep highest value, and 3) append saved fndds and ferming code to new df 

for (i in 1:row_count){
  #store the fndds code
  fndds_code <- FNDDS_ferm_bin1$parentfoodcode[i]
  
  # filter df down to the specific fndds code
  fndds_info <- FNDDS_ferm_ingcode %>%
    filter(parentfoodcode == fndds_code)
  
  # count number of fndds rows in saved df
  fndds_entries <- fndds_info %>%
    nrow()
  
  # if fndds temp df has more than one row, do the comparison. Else move on
  if (fndds_entries != 1) {
      
      for (j in 1:fndds_entries){
        
        # if temp variable is less than ferming code at row j
        if(temp < fndds_info$ingfermscore[j]){
          
          # overwrite temp with new ferming
          temp <- fndds_info$ingfermscore[j]
          
          # else, move on to next loop
          } else {
            next 
          }
      }
   # else, do nothing BUT finish loop code
  } else {
     
  }
  
  # save ferming and fnnds value to df
  temp_ferming <- data.frame(parentfoodcode=fndds_code,
                           merged_ferming=temp)
  
  # reset temp 
  temp <- 0

  # append i and j values to blank data frame
  FNDDS_ferm_ingcode_merged <- bind_rows(FNDDS_ferm_ingcode_merged ,temp_ferming)
}

# merge ffbin1 and ferm ing info to one df
FNDDS_ferm_info <- inner_join(FNDDS_ferm_ingcode_merged, FNDDS_ferm_bin1[, c("ffbin1", "ffcont1", "ffbin9","ffcont9", "lmcat", "parentfoodcode")], by = "parentfoodcode")

# clean up environment 
rm(FNDDS_ferm_bin1)
rm(FNDDS_ferm_ingcode)
rm(FNDDS_ferm_ingcode_merged)
rm(FNDDS_ferm_zpg)
rm(fndds_info)
rm(temp_ferming)
rm(fndds_code)
rm(fndds_entries)
rm(i)
rm(j)
rm(row_count)
rm(temp)
```

```{r Subset Ingredients and Master FNDDS files}
library(stringmagic) # for paste with operation chaining 

# this version retains the Seq_Num and Ingredient_code and Ingredient_description columns for the different formulations of items with the same FNDDS code.

# adjust df and subset columns to just the ingredient relevant ones
FNDDS_ingredients <- FNDDS_desc_merged %>% 
  ungroup() %>%
  rename(Food_code_key = Food_code) %>%
  select(Food_code_key, Ingredient_description, Ingredient_code, Seq_num) %>%
  arrange(Food_code_key, Seq_num) %>%
  mutate(Ingredient_key = row_number(), .after = 0) %>%
  mutate(cto_key = Seq_num + (Food_code_key*100), .after = 4)

# merge rows with the same FNDDS codes and remove unnecessary seq-num column
FNDDS_collapsed <- FNDDS_desc_merged %>% 
  group_by(WWEIA_Category_number, WWEIA_Category_1, Category_1_code, WWEIA_Category_2, 
           Category_2_code, WWEIA_Category_3, Category_3_code, Main_food_description, Additional_food_description,
           Condiment_status, Mixdish_status,Food_code) %>%
  summarise(Seq_num = 
              paste(na.omit(Seq_num), collapse = "***")) %>%
  mutate(cto_search = paste(Main_food_description, Additional_food_description, sep = " ||| ")) %>%
  mutate(Seq_num = NULL)

# number all the rows in main df and rename
FNDDS_collapsed <- FNDDS_collapsed %>%
  ungroup() %>%
  rename(Food_code_key = Food_code)

# manipulate df columns as needed
FNDDS_ferm_info <- FNDDS_ferm_info %>%
  rename(Food_code_key = parentfoodcode) %>%
  mutate(Food_code_key = as.numeric(Food_code_key))

# add the ferm information from ZPG's analysis to the dataset
FNDDS_collapsed <- inner_join (FNDDS_collapsed, FNDDS_ferm_info, by = "Food_code_key")
```

After finishing the fuzzy search compatible file, the next step is to do the same for the portion sizes linked to FNDDS codes
```{r Subset Portions file}
library(snakecase) # adjust column names
 
# The FNDDS 21-23 file portion options and information for each FNDDS code
FNDDS_portions <- read_wides(files = "FNDDS_portions_raw.csv")

# rename and remove columns 
FNDDS_portions$file <- NULL

colnames(FNDDS_portions) <- to_snake_case(colnames(FNDDS_portions))

# manipulating columns as needed 
FNDDS_portions <- FNDDS_portions %>%
  rename(food_code_key = food_code) %>%
  mutate(portion_key = row_number(), .after = 5)
```

```{r checking labels alignment}

# count Mixdish 1s by category
FNDDS_collapsed %>%
  group_by(WWEIA_Category_1) %>%
  count(Mixdish_status)

# count Condi 1s by category
FNDDS_collapsed %>%
  group_by(WWEIA_Category_1) %>%
  count(Condiment_status)
```

My previous method for adding inulin status 1 or 0 ended up with a misaligned df and caused duplicate rows to pop up (5438 instead of 5432). Because I want to move onto CTO coding, I've decided to add inulin status after everything has been condensed by FNDDS code to hopefully prevent that issue. It would probably be worthwhile to fix my original method at some point (possible just for practice at that point).

```{r Adding inulin status by category }
#pull out categories that will be labeled 1 for inulin
inulin_values <- Item_values %>%
  filter(Inulin_status == 1)

#count total categories
wweia_amount <- Item_values %>%
  nrow()

# count inulin categories
inulin_amount <- inulin_values %>%
  nrow()

# blank df to copy later
inulin_holder <- data.frame()

# The following code is repeated for all the wweia categories. For the number of whole categories that need the inulin label 1) save the CTO value for the category and for each wweia category 2) check if they are the same or not. If equal, filter the FNDDS codes to that category, change the inulin status to 1, and add those rows to a holder df. If they are not equal, move on to the next category.

# Category 1 codes
for(i in 1:inulin_amount) {
  wweia_cat <- inulin_values$Category_key[i]
  
  for(j in 1:wweia_amount){
    
    if(j == wweia_cat) {
      category_holder <- FNDDS_collapsed %>%
        filter(Category_1_code == wweia_cat) %>%
        mutate(Inulin_status = '1') %>%
        select(Food_code_key, Category_1_code, Inulin_status)
      
      inulin_holder <- bind_rows(inulin_holder, category_holder)
    } else {
      next
    }
  }
}

# Category 2 code

for(i in 1:inulin_amount) {
  wweia_cat <- inulin_values$Category_key[i]
  
  for(j in 1:wweia_amount){
    
    if(j == wweia_cat) {
      category_holder <- FNDDS_collapsed %>%
        filter(Category_2_code == wweia_cat) %>%
        mutate(Inulin_status = '1') %>%
        select(Food_code_key, Category_1_code, Inulin_status)
      
      inulin_holder <- bind_rows(inulin_holder, category_holder)
    } else {
      next
    }
  }
}

# Category 3 code

for(i in 1:inulin_amount) {
  wweia_cat <- inulin_values$Category_key[i]
  
  for(j in 1:wweia_amount){
    
    if(j == wweia_cat) {
      category_holder <- FNDDS_collapsed %>%
        filter(Category_3_code == wweia_cat) %>%
        mutate(Inulin_status = '1') %>%
        select(Food_code_key, Category_1_code, Inulin_status)
      
      inulin_holder <- bind_rows(inulin_holder, category_holder)
    } else {
      next
    }
  }
}

# Manual Codes: after doing a manual check, here are additional codes that should be flagged. JOC has a separate word file that contains more information about why each code was flagged.

inulin_manual <- c(72116000, 92202010, 92152000, 95312400, 95313200, 93102000, 12220200, 12220270, 12220280, 32103045, 32103035, 32103015, 32103025, 27446255, 27446245, 27446225, 27446235, 27450067, 27450065, 27450061, 27450063, 58148115, 58148117, 58148111, 58148113, 71601045, 71601035, 71601015, 71601025, 71603045, 71603035 ,71603015, 71603025, 92309030, 92309020, 92309510, 92305110, 92305090, 92305920, 83204000, 83204050, 83201000, 83203000, 83202020, 83204500, 83205450, 83200100, 83206500, 83207000, 91301081, 91301510, 11830165)

for(i in inulin_manual) {
  category_holder <- FNDDS_collapsed %>%
    filter(Food_code_key == i) %>%
    mutate(Inulin_status = '1') %>%
    select(Food_code_key, Category_1_code, Inulin_status)
      
    inulin_holder <- bind_rows(inulin_holder, category_holder)
}

# collapse duplicate fndds codes
inulin_collapsed <- inulin_holder %>%
  mutate(Inulin_status = as.numeric(Inulin_status)) %>%
  group_by(Food_code_key) %>%
  summarise(Inulin_status = mean(Inulin_status))

FNDDS_collapsed <- merge(FNDDS_collapsed, inulin_collapsed, by = "Food_code_key", all.x = TRUE)

FNDDS_collapsed$Inulin_status[is.na(FNDDS_collapsed$Inulin_status)] <- 0

FNDDS_collapsed <- FNDDS_collapsed %>%
  relocate(Inulin_status, .after = Mixdish_status)

#clear environment
rm(category_holder)
rm(inulin_collapsed)
rm(inulin_holder)
rm(inulin_values)
rm(i)
rm(j)
rm(inulin_amount)
rm(inulin_manual)
rm(wweia_amount)
rm(wweia_cat)
```


26.08.05 Some preliminary tested showed I needed to add additional WWEIA labels into mixed dish type for CTO. They are all in cat 3, so I need separate code to make it work.

```{r Category 3 Mix Dish Status}
library(tidyverse) # general
library(dplyr) # piping format

# create df with only category 3 values
category_level_values <- Item_values %>%
  filter(Grouping == "Category 3")

Group_value_holder <- FNDDS_desc_merged[,c('Category_3_code','Food_code')]

# get the number of categories for level 3
row_count <- Item_values %>%
  filter(Grouping == "Category 3") %>%
  nrow()

# blank data frame to use in the for loop
Item_codes_Merged <- data.frame() 

# loop to go through each level group and add the survey CTO code value
for (i in 1:row_count) {
  # load category into variable
  current_group <- category_level_values$Category_key[i]
  
  # load mix dish status into variable
  mixdish_status <- category_level_values$Mixdish_status[i]
  
  if(mixdish_status == 1){
    # second df to hold info. Filters Group_value_holder by WWEIA category before adding columns with pasted CTO code value and condiment status 
      Group_value_holder_2 <- Group_value_holder %>% filter(Category_3_code == current_group) %>%
        mutate(Cat_3_mixdish = mixdish_status) %>%
        select(c(Food_code, Cat_3_mixdish))
      
      # adds newly saved CTO values to passed df
      Item_codes_Merged <- bind_rows(Item_codes_Merged,Group_value_holder_2)
      
      # prints current group and total row count of passed df
      print(paste0(current_group,":",count(Item_codes_Merged)))
  } else {
    next
  }
  
}

# insert inulin values by FNDDS code
FNDDS_collapsed <- merge(FNDDS_collapsed, unique(Item_codes_Merged), by.x = "Food_code_key",by.y = "Food_code", all.x = TRUE)

# clear out 0's from mixdish column
FNDDS_collapsed$Mixdish_status[FNDDS_collapsed$Mixdish_status == 0] <- NA

# merge columns, replace NA with 0
FNDDS_collapsed <- FNDDS_collapsed %>% 
  mutate(Mixdish_status = coalesce(Mixdish_status,Cat_3_mixdish))

FNDDS_collapsed$Mixdish_status[is.na(FNDDS_collapsed$Mixdish_status)] <- 0

# remove unnecessary column
FNDDS_collapsed$Cat_3_mixdish <- NULL

# order df correctly
FNDDS_collapsed <- FNDDS_collapsed %>% 
  arrange(WWEIA_Category_number)

# clean up environment
rm(category_level_values)
rm(Group_value_holder)
rm(Group_value_holder_2)
rm(Item_codes_Merged)
rm(Item_values)
rm(current_group)
rm(mixdish_status)
rm(i)
rm(row_count)
```

### End Results 

Final csv files for now.

```{r Searchable FNDDS codes for CTO}
# save collapsed df csv 
write.csv(FNDDS_collapsed,"C:/Users/josecoll/Downloads/2026.07.10 FNDDS Files/fndds23_master_JOC.csv", row.names = FALSE)

# save ingredients as a csv 
write.csv(FNDDS_ingredients,"C:/Users/josecoll/Downloads/2026.07.10 FNDDS Files/fndds23_ingredients_JOC.csv", row.names = FALSE)

# save edited df as csv for survey CTO 
write.csv(FNDDS_portions,"C:/Users/josecoll/Downloads/2026.07.10 FNDDS Files/fndds23_portions_JOC.csv", row.names = FALSE)

```
